---
title: "Data180_final"
output: html_document
date: "2023-12-10"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Read in csv file 
```{r}
loan_default <- read.csv("loan_default_data_set.csv")
```
## R Markdown

1. Data wrangling
a. What is the dimension (shape) of the dataset?  How many rows and columns does the data set have?

```{r}
dim(loan_default)
```
This data set contains 20,000 rows and 21 columns


b. Report the column names of the data set.
```{r}
colnames(loan_default)
```

c. Which types of data are there in the dataset? Numeric, categorical, ordinal?
This dataset contains numerical and categorical data. 

d. Which columns contain missing values and how much (what percent) of those columns are missing?
```{r}
sapply(loan_default, function(x) which(is.na(x)))
sapply(loan_default, function(x) sum(is.na(x)))
#percent missing pct_card_over_50_uti
(1958/20000)*100= 9.79
#percent missing rep_income
(1559/20000)*100= 7.795
```
Column "pct_card_over_50_uti" contains 1958 missing values (9.79% of total) and "rep_income" contains 1559 missing values (7.795% of total). 

e. How do you think we should deal with missing values? 
Missing values can be dealt with in a variety of ways. The easiest way would be to drop the missing values from the dataset

f.
g. 
```{r}

```